On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-08-19 paper.tex
Writing section on Rehmann theorem

Let $\Phi$ be a reduced irreducible root system of rank $\geq 3$.
The aim of this section is to find a presentation of the Steinberg group $\St(\Phi, B)$ in terms of generators $x_{\alpha}(\xi)$
 for which $\xi$ is homogeneous of degree $\leq 1$, see~\cref{prop:tul3.3} below.
Our presentation should be thought of as an analogue of~\cite[Lemma~3.3]{Tu83} in the case $n=4$ (with variables $t$ and $t^{-1}$ swapped).
Notice that Tulenbaev's short proof of \cite[Lemma~3.3]{Tu83} which works for $\Phi=\rA_\ell$, $\ell\geq 4$ does not generalize to the case $\ell=3$, 
 so we are forced to recourse to a much longer argument inspired by the proof of Rehmann--Soul{\'e} theorem on
 finite presentation of Steinberg groups over polynomial rings (cf.~\cite{Re75, RS76}).

@@

\begin{rem}
There are several reasons why we can not simply refer to~\cite{Re75} or~\cite{RS76} and need to reproduce arguments from these papers.
\begin{itemize}
 \item Rehmann and Soul{\'e} deal with usual polynomial rings rather than $\mathbb{Z}$-graded rings
  (i.\,e. their results correspond to the special case $\mathfrak{m}=0$ of our result);
 \item Presentations from~\cite{Re75,RS76} are formulated in terms of generating set $\XX{2}$ which is not sufficient for our purposes
  (we need a presentation in terms of $\XX{1}$);
 \item it is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}, while we work with
  arbitrary coefficient ring $A$.
\end{itemize} 
\end{rem}

For $m\geq 2$ define $G_m$ as the group presented by generators $\mathcal{X}_m$ and the following three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_{\alpha}(\xi) x_{\alpha}(\eta) & = x_{\alpha}(\xi+\eta),&  \xi,\eta\in B_d,\ d\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\beta}(\eta))] &  = 1, & \text{in the case $\alpha+\beta\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_d,\ \eta \in B_e,\ d,e\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\beta}(\eta)] & = x_{\alpha+\beta}(N_{\alpha,\beta}\xi\eta), & \text{in the case $\alpha+\beta\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_d,\ \eta \in B_e,\ d, e, d+e \leq m;
 \end{align}
For $m=1$ define $G_1$ as the group presented by generators $\mathcal{X}_m$, three families of relations $a_1, b_1, c_1$ as above and the following extra family of relations:
\begin{align}  \label{eq:d1} \tag{$d_1$} [[x_{\alpha}(\xi), x_{\beta}(\eta)], x_{\gamma}(\zeta) ] & = 1, & \xi,\eta,\zeta\in B_1,\ \alpha + \beta \in \Phi,\ \alpha + \beta + \gamma \not\in \Phi\cup\{0\}.
\end{align}
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

It is clear that the natural embedding of generators $\mathcal{X}_n \subseteq \mathcal{X}_{n+1}$ induces
 a map $\theta_n\colon G_n \to G_{n+1}$.
Our main goal is to show the following result.
\begin{prop}\label{prop:tul3.3}
 For $n\geq 1$ the map $\theta_n$ is an isomorphism. Consequently, $\St(\Phi, B)$ can be presented by generators $\mathcal{X}_1$
  and four families of relations $a_1$, $b_1$, $c_1$, $d_1$.
\end{prop}

 \begin{lemma} In the group $G_1$ the following relation holds:
\begin{align} 
 [[x_{\alpha}(a t), x_{\beta}(b t)],  x_{\gamma}(c)] & = [x_{\alpha+\beta}(a b t),  x_{\gamma}(c t)].
\end{align}
 \end{lemma}

%Denote by $F(S)$ the free group on the generating set $S$.
%It is clear that $\RR{n}\subseteq\RR{n+1}$, moreover $\langle \RR{n} \rangle = \langle \RR{\infty}\rangle \cap F(\XX{n})$.
%well, not so clear, actually :)
To prove~\cref{lem:tul3.3} it suffices to show that the natural map $\theta_m \colon G_m \to G_{m+1}$ is an isomorphism for $m\geq 2$.
Indeed, this would give the required isomorphism $G_1 \cong \St(\Phi, B)$ since $\langle \XX{\infty} \mid \RR{\infty} \rangle$ is evidently isomorphic to $\St(\Phi, B)$.

@@

To prove that $\theta_m$ is an isomorphism we construct the missing generators of $G_{m+1}$ in $G_m$ and then show that they satisfy relations $\RR{m+1}$.

@@

To verify relations $(c_{m+1})$ it suffices to show that
\begin{equation}
\label{eq:verify-bmp1} [x_\alpha(\xi), x_{\alpha'}(at^{m+1})] = [x_\alpha(t\xi), x_{\alpha'}(at^m)],\ a \in A,\ \xi \in B_k,\ k\leq 0.
\end{equation}
We can find root subsystem $\Psi \subseteq \Phi$ of type $\mathsf{A}_3$ containing the roots $\alpha, \alpha'$.
Choose basis $\{\alpha,\beta,\gamma\}$ of $\Psi$ so that $\alpha'=\beta$ and the Dynkin diagram of $\Psi$ looks as follows:
\begin{equation}\label{eq:diagA3} \xymatrix{\circ_\alpha \ar@{-}[r] & \circ_\beta \ar@{-}[r] & \circ_\gamma}.\end{equation}
Let us verify~\eqref{eq:verify-bmp1}:
\begin{align*}
   [x_\alpha(\xi), x_\beta(at^{m+1})] = [x_\alpha(\xi), [x_{\beta + \gamma}(t), x_{-\gamma}(a't^m)]]&  \text{ by~\eqref{eq:defmpn} for a suitable $a'\in A$} \\ 
 = [x_{\alpha+\beta+\gamma}(\epsilon t\xi), {}^{x_{\beta+\gamma}(t)}x_{-\gamma}(a't^m)]             &  \text{ by~\eqref{eq:H1iii}, for $\epsilon=N_{\alpha, \beta+\gamma}$} \\
 = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t\xi), x_{-\gamma}(a't^m)]             &  \text{ by $(b_1)$} \\
 = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t^2\xi), x_{-\gamma}(a't^{m-1})]       &  \text{ by~\eqref{item:i} if $k=0$ or~\eqref{eq:cm} if $k <0$} \\
 %TODO: There's a mistake at this moment --- k = 0, m = 1 and the needed relation is not in G_1 -- :(
 = [[x_\alpha(t\xi), x_{\beta+\gamma}(t)], {}^{x_{\beta+\gamma}(t)} x_{-\gamma}(a't^{m-1})]         &  \text{ by~$(b_2)$, $(c_2)$ or by~\eqref{item:ii},\eqref{eq:defmpn} if $m=1$} \\
 = [x_\alpha(t\xi), x_\beta(at^m)]                                                                  &  \text{ by~\eqref{eq:H1iii}.}
\end{align*}

>>>>

Let $\Phi$ be a reduced irreducible root system of rank $\geq 3$.
The aim of this section is to find a presentation of the Steinberg group $\St(\Phi, B)$ in terms of generators $x_{\alpha}(\xi)$
 for which $\xi$ is homogeneous of degree $\leq 1$, see~\cref{prop:tul3.3} below.
Our presentation should be thought of as an analogue of~\cite[Lemma~3.3]{Tu83} in the case $n=4$ (with variables $t$ and $t^{-1}$ swapped).
Notice that Tulenbaev's short proof of \cite[Lemma~3.3]{Tu83} which works for $\Phi=\rA_\ell$, $\ell\geq 4$ does not generalize to the case $\ell=3$,
%TODO: Explain why we are dealing with general \Phi rather than just the case \rA_3 -- Reason the difficulty is the same
 so we are forced to recourse to a much longer argument inspired by the proof of Rehmann--Soul{\'e} theorem on
 finite presentation of Steinberg groups over polynomial rings (cf.~\cite{Re75, RS76}).

We need one additional technical definition. We call an ordered triple of roots $(\alpha, \beta, \gamma) \in \Phi \times \Phi \times \Phi$ an {\it $\rA_3$-triple}
 if these vectors form a basis of a root subsystem $\Psi \subseteq \Phi$ of type $\rA_3$, and moreover if the Dynkin diagram of $\Psi$ with respect to this basis has the form
 $\xymatrix{\circ_\alpha \ar@{-}[r] & \circ_\beta \ar@{-}[r] & \circ_\gamma}.$ %TODO: Draw better diagram :)

@@

\begin{comment} 
\begin{rem}
There are several reasons why we can not simply refer to~\cite{Re75} or~\cite{RS76} and need to reproduce arguments from these papers.
\begin{itemize}
 \item Rehmann and Soul{\'e} deal with usual polynomial rings rather than $\mathbb{Z}$-graded rings
  (i.\,e. their results correspond to the special case $\mathfrak{m}=0$ of our result);
 \item Presentations from~\cite{Re75,RS76} are formulated in terms of generating set $\XX{2}$ which is not sufficient for our purposes
  (we need a presentation in terms of $\XX{1}$);
 \item It is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}, while we work with
  arbitrary coefficient rings $A$.
\end{itemize} 
\end{rem}
\end{comment} %TODO: Find proper place for this discussion - it may not be appropriate here

For $m\geq 2$ define $G_m$ as the group presented by generators $\mathcal{X}_m$ and the following three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_{\alpha}(\xi) x_{\alpha}(\eta) & = x_{\alpha}(\xi+\eta),&  \xi,\eta\in B_d,\ d\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\beta}(\eta))] &  = 1, & \text{in the case $\alpha+\beta\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_d,\ \eta \in B_e,\ d,e\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\beta}(\eta)] & = x_{\alpha+\beta}(N_{\alpha,\beta}\xi\eta), & \text{in the case $\alpha+\beta\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_d,\ \eta \in B_e,\ d, e, d+e \leq m.
 \end{align}
For $m=1$ define $G_1$ as the group presented by generators $\mathcal{X}_m$, three families of relations $a_1, b_1, c_1$ as above and the following additional family of relations:
\begin{align}  \label{eq:d1} \tag{$d_1$} [[x_{\alpha}(\xi), x_{\beta+\gamma}(\eta)], x_{\beta}(\zeta) ] & = 1, 
 & \xi,\eta,\zeta\in B_1,\ \text{ $(\alpha, \beta, \gamma)$ is an $\rA_3$-triple }.
\end{align}
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

It is clear that the natural embedding of generators $\mathcal{X}_n \subseteq \mathcal{X}_{n+1}$ induces
 a map $\theta_n\colon G_n \to G_{n+1}$.
Our main goal is to show the following result.
\begin{prop}\label{prop:tul3.3}
 For $n\geq 1$ the map $\theta_n$ is an isomorphism. Consequently, $\St(\Phi, B)$ can be presented by generators $\mathcal{X}_1$
  and families of relations $(a_1)$, $(b_1)$, $(c_1)$, $(d_1)$.
\end{prop}
%TODO: Deduce 2nd assertion from 1st
To prove~\cref{lem:tul3.3} it suffices to show that the natural map $\theta_m \colon G_m \to G_{m+1}$ is an isomorphism for $m\geq 2$.
Indeed, this would give the required isomorphism $G_1 \cong \St(\Phi, B)$ since $\langle \XX{\infty} \mid \RR{\infty} \rangle$ is evidently isomorphic to $\St(\Phi, B)$.
%Denote by $F(S)$ the free group on the generating set $S$.
%It is clear that $\RR{n}\subseteq\RR{n+1}$, moreover $\langle \RR{n} \rangle = \langle \RR{\infty}\rangle \cap F(\XX{n})$.
%well, not so clear, actually :)

@@

\begin{lemma} For every $\rA_3$-triple $(\alpha, \beta, \gamma)$ in $\Phi$ the following relation holds in the group $G_1$:
\begin{align} 
 [[x_{\beta}(b t), x_{\alpha}(a t)],  x_{\gamma}(c)] & = [x_{\alpha+\beta}(a b t),  x_{\gamma}(c t)].
\end{align}
 \end{lemma}
\begin{proof}
 \begin{align*}
   [[x_{\beta}(b t), x_{\alpha}(a t)],  x_{\gamma}(c)] = [{}^{x_{\beta}(b t)}\!x_{\alpha}(at), [x_{\gamma}(c),x_{\beta}(b t)]] &  \text{ by~\eqref{eq:H1iii} and~$(b_1)$} \\ 
                                                       = {}^{x_{\beta}(b t)}[ x_{\alpha}(at), x_{\beta +\gamma}(bct) ] & \text{ by $(b_1)$ and $(c_1)$}\\
                                                       = [x_{\alpha}(at), x_{\beta +\gamma}(bct)] & \text{ by $(d_1)$. \qedhere}
 \end{align*}

\end{proof}

To prove that $\theta_m$ is an isomorphism we construct the missing generators of $G_{m+1}$ in $G_m$ and then show that they satisfy relations $a_{m+1}$, $b_{m+1}$ and $c_{m+1}$.

@@

To verify relations $(c_{m+1})$ it suffices to show that
\begin{equation}
\label{eq:verify-bmp1} [x_\alpha(\xi), x_{\alpha'}(at^{m+1})] = [x_\alpha(t\xi), x_{\alpha'}(at^m)],\ a \in A,\ \xi \in B_k,\ k\leq 0.
\end{equation}
We can find root subsystem $\Psi \subseteq \Phi$ of type $\mathsf{A}_3$ containing the roots $\alpha, \alpha'$.
Choose basis $\{\alpha,\beta,\gamma\}$ of $\Psi$ so that $\alpha'=\beta$ and the Dynkin diagram of $\Psi$ looks as follows: HERE WAS A DIAGRAM

Let us verify~\eqref{eq:verify-bmp1}:
\begin{align*}
   [x_\alpha(\xi), x_\beta(at^{m+1})] = [x_\alpha(\xi), [x_{\beta + \gamma}(t), x_{-\gamma}(a't^m)]]&  \text{ by~\eqref{eq:defmpn} for a suitable $a'\in A$} \\ 
 = [x_{\alpha+\beta+\gamma}(\epsilon t\xi), {}^{x_{\beta+\gamma}(t)}x_{-\gamma}(a't^m)]             &  \text{ by~\eqref{eq:H1iii}, for $\epsilon=N_{\alpha, \beta+\gamma}$} \\
 = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t\xi), x_{-\gamma}(a't^m)]             &  \text{ by $(b_1)$} \\
 = {}^{x_{\beta+\gamma}(t)}[x_{\alpha+\beta+\gamma}(\epsilon t^2\xi), x_{-\gamma}(a't^{m-1})]       &  \text{ by~\eqref{item:i} if $k=0$ or~\eqref{eq:cm} if $k <0$} \\
 %TODO: There's a mistake at this moment --- k = 0, m = 1 and the needed relation is not in G_1 -- :(
 = [[x_\alpha(t\xi), x_{\beta+\gamma}(t)], {}^{x_{\beta+\gamma}(t)} x_{-\gamma}(a't^{m-1})]         &  \text{ by~$(b_2)$, $(c_2)$ or by~\eqref{item:ii},\eqref{eq:defmpn} if $m=1$} \\
 = [x_\alpha(t\xi), x_\beta(at^m)]                                                                  &  \text{ by~\eqref{eq:H1iii}.}
\end{align*}
